### 2019-02-23 manuscript/introduction.tex
- adding .gitignore to remove nuisance files (e.g. accidental Rhistory editing) - modified introduction - modified addresses to conform to NP journal standards

Species accumulate around the tree of life at different rates.
The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016}. % todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky
In these studies, it is common to identify a single focal trait and investigate its association with rates of speciation and extinction. % (but see... ?) % refs: there must be some papers that musse on multiple traits? especially for geographic range?
This is problematic for two reasons.
On the biology side, the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
On the methods side, excluding other traits that additionally influence diversification can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
One approach to this problem is to allow for a `hidden' or unspecified trait in the analysis \citep{beaulieu_2016}. % and an earlier beaulieu? applications of hisse?
This can be effective, but it leaves behind the question of whether the hidden trait corresponds to a real trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
Alternatively, we can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
% B: We can then have a good prior, reliable data, and replicate the studies.  E: Agree. But when I tried to add it, I realized hidden models can also be hypothesis-generating, with the purported hidden trait identified in the next clade.
In this paper, we focus on two of the best studied traits in flowering plants, ploidy and breeding system, and investigate their individual and joint effects on diversification.

Polyploidization multiplies the genomic content of cells, and it consequently has the potential to affect many or all traits and a great variety of evolutionary processes.
It is also a mutation that occurs commonly in plants, at it is widespread at both population and lineage scales.
The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century \citep{stebbins1938}. % B: instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout. E: What word would that be? I was taking "ploidy" to mean "whatever multiple," but is that a terminological abuse?
The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent diploidization, across angiosperms \citep{lynch2000, vision2000}. % E: I am deferring to R's use of "diploidization" instead of "re-diploidization"
Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. % todo refs, jiao etc
%
The prevalence of polyploidy, its variation across clades, and its large effects on genotypes and phenotypes raises the hypothesis that ploidy plays an important role in shaping rates of speciation and extinction.
This has led to a debate: \citet{mayrose_2011, mayrose_2015} found a lower rate of lineage diversification for polyploids, while \citet{soltis_2014} argued that polyploidy is not an evolutionary dead end because whole genome duplications occurred at the base of many successful angiosperm clades, and \citet{landis_2018} reported higher diversification rates following some whole genome duplications. % and tank2015?
Our study revisits this question, additionally incorporating the process of diploidization.
% E:  Rosana, could you add a little about the importance of including diploidization?  todo
Furthermore, we consider the interaction between polyploidy and other traits.

From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats. % todo refs: husband etc
Other correlations may have a more direct, causal connection.
Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
% barrett2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system. However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
This creates a correlation between polyploidy and self-compatibility by preventing the existence of self-incompatible polyploids.
In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.

>>>>

Species accumulate around the tree of life at different rates.
The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016}. % todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky
In these studies, it is common to identify a single focal trait and investigate its association with rates of speciation and extinction. % (but see... ?) % refs: there must be some papers that musse on multiple traits? especially for geographic range?

%B: I understand the apparent contrast, but biology vs. methods is not a properly framed Venn diagram, IMO. %B: was: On the biology side, the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
%This is problematic for two reasons.
This is problematic because the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
%B: was: On the methods side, excluding other traits that additionally influence diversification can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
One approach to this problem is to allow for a `hidden' or unspecified trait in the analysis \citep{beaulieu_2016}. % and an earlier beaulieu? applications of hisse? %B: perhaps above can begin with: "One proposed approach to accommodate..." 
This can be effective, but it leaves behind the question of whether the hidden trait corresponds to a real trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
Alternatively, we can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
% B: We can then have a good prior, reliable data, and replicate the studies.  E: Agree. But when I tried to add it, I realized hidden models can also be hypothesis-generating, with the purported hidden trait identified in the next clade. %B: I don't get "the next clade"?
In this paper, we focus on two of the best studied traits in flowering plants, ploidy and breeding system, and investigate their individual and joint effects on diversification.

Polyploidization multiplies the genomic content of cells, and it consequently has the potential to affect many or all traits and a great variety of evolutionary processes.
It is also a mutation that occurs commonly in plants, at it is widespread at both population and lineage scales.
The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century \citep{stebbins1938}. % B: instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout. E: What word would that be? I was taking "ploidy" to mean "whatever multiple," but is that a terminological abuse? %B: sorry, that was supposed to complain about the use of "polyploidy" instead of "ploidy" to designate the trait. The trait is ploidy, and a state is polyploidy. Also, we should clarify that polyploidy is any of tetrapoidy+. I am not sure what we did with 3x here. It's rather rare, but...?
The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent diploidization, across angiosperms \citep{lynch2000, vision2000}.
Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. % todo refs, jiao etc
%
The prevalence of polyploidy, its variation across clades, and its large effects on genotypes and phenotypes raises the hypothesis that ploidy plays an important role in shaping rates of speciation and extinction.
This has led to a debate: \citet{mayrose_2011, mayrose_2015} found a lower rate of lineage diversification for polyploids, while \citet{soltis_2014} argued that polyploidy is not an evolutionary dead end because whole genome duplications occurred at the base of many successful angiosperm clades, and \citet{landis_2018} reported higher diversification rates following some whole genome duplications. % and tank2015?
Our study revisits this question, additionally incorporating the process of diploidization.
% E:  Rosana, could you add a little about the importance of including diploidization?  todo
Furthermore, we consider the interaction between polyploidy and other traits.

From the beginning, study of this variable cytogenetic property considered correlations with other traits \citep{stebbins1938}. %B: ploidy research, or nothing (as now); was: "From the very beginning of polyploidy research," polyploidy is a state of the trait.
Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats. % todo refs: husband etc
Other correlations may have a more direct, causal connection.
Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950}. %B: removed barrett1988, because he highlights correlation between polyploidy and self-compatibility, not self-fertilization.
In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
% barrett2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system. However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
This creates a correlation between polyploidy and self-compatibility by precluding the existence of self-incompatible polyploids. %B: prevent twice in a row; replaced with preclude.
In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.

### 2019-02-23 manuscript/main.tex
- new paragraph(s) in introduction, and matching citations - new title ideas or modifications
- adding .gitignore to remove nuisance files (e.g. accidental Rhistory editing) - modified introduction - modified addresses to conform to NP journal standards

% Accumulate ideas for titles here...
%
% When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae % E: Fun, but I think we need something more formal
%
% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
%
% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal

@@

\noindent
Rosana Zenil-Ferguson,%
\endnote{University of Minnesota Twin Cities}
%
\noindent
J.~Gordon Burleigh,%
\endnote{University of Florida}
%
\noindent
William A.~Freyman,%
\endnote{23 and me}
%
\noindent
Boris Igi\'c,%
\endnote{University of Illinois}
%
\noindent
Itay Mayrose,%
\endnote{Tel Aviv University}
%
and
Emma E.~Goldberg%
\endnote{University of Minnesota Twin Cities}

@@

\noindent
Author for correspondence: Rosana Zenil-Ferguson

>>>>

% Accumulate ideas for titles here...
%
% When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae % E: Fun, but I think we need something more formal
%
%% When the favorite trait is not enough: the link of polyploidy and breeding system in diversification %B: <- eh?
%
% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
%
% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
%
%% Interaction of ploidy and breeding system in lineage diversification %B: <- how 'bout that?

@@

%B: I included extended affiliations necessary at submission time. Check, please! 
%B->R you may wish to add new affiliation at Manoa, but I left the original state alone.
\noindent
Rosana Zenil-Ferguson\endnote{Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.}$^{\dagger}$,
%
\noindent
J.~Gordon Burleigh\endnote{Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.},
%
\noindent
William A.~Freyman\endnote{23andMe, Inc., 899 W Evelyn Ave, Mountain View, CA 94041, U.S.A.},
%
\noindent
Boris Igi\'c\endnote{Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. },
%
\noindent
Itay Mayrose\endnote{Department of Molecular Biology and Ecology of Plants, Tel Aviv University, Tel Aviv, Israel},
%
and
Emma E.~Goldberg\endnote{Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.}
%
\vfill

@@

\noindent
$\dagger$Author for correspondence.
%Rosana Zenil-Ferguson
\vfill
